\textbf{Unshuffling Permutations}

### 2016-04-06 slides/slides.tex
Restructuration (pas finie)

@@

\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{std}}
\DeclareMathOperator{\SHUFFLE}{\bullet}

@@

\date[LATIN 16]{
    {
    \scriptsize Latin American Theoretical Informatics Symposium.\\
    April $11$-$15$ $2016$
    }
}% end date

@@

@@

@@

\begin{block}{Proposition}
    The set of binary words of length $n$ that are square w.r.t.
    $\shuffle$ is in one-to-one correspondence with the set of
    square permutations of length $n$ avoiding the patterns
    $213$ and~$231$.
\end{block}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{}

>>>>

% Affichage du plan à chaque changement de partie.
\AtBeginSection[]{
    \frame<handout:0>{
        \frametitle{Plan}
        \tableofcontents[
            sectionstyle=show/hide,
            subsectionstyle=show/shaded/hide]}
}

@@

\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{std}}
\DeclareMathOperator{\SHUFFLE}{\bullet}
\DeclareMathOperator{\La}{\mathtt{a}}
\DeclareMathOperator{\Lb}{\mathtt{b}}

@@

\date[LATIN 16]{
    {
    \scriptsize Latin American Theoretical Informatics Symposium \\
    April $11$-$15$ $2016$
    }
}% end date

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Contents}
    \tableofcontents
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Shuffling words}
\begin{block}{Definition}
    The \alert{shuffle product} $\shuffle$ on words of $A^*$ is defined
    inductively by
    \begin{align*}
        u \shuffle \epsilon &= \epsilon \shuffle u = \{u\} \\
        (ua \shuffle vb) &= (ua \shuffle v)b \cup (u \shuffle vb)a
    \end{align*}
\end{block}
\medskip

\uncover<2->{%
To take into account multiplicities, we consider $\shuffle$ as a linear
product
\begin{equation*}
    \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
\end{equation*}
on $\QQ[A^*]$, the $\QQ$-linear span of words defined by
\begin{align*}
    u \shuffle \epsilon & = \epsilon \shuffle u = u \\
    (ua \shuffle vb) &= (ua \shuffle v)b + (u \shuffle vb)a
\end{align*}}

\uncover<3->{%
\begin{block}{Example}
    \vspace{-1em}
    \begin{equation*}\begin{split}
        \La\Lb \shuffle \textcolor{BrickRed}{\Lb\La} & =
        \La\Lb\textcolor{BrickRed}{\Lb\La}
        + \La\textcolor{BrickRed}{\Lb}\Lb\textcolor{BrickRed}{\La}
        + \La\textcolor{BrickRed}{\Lb\La}\Lb
        + \textcolor{BrickRed}{\Lb}\La\Lb\textcolor{BrickRed}{\La}
        + \textcolor{BrickRed}{\Lb}\La\textcolor{BrickRed}{\La}\Lb
        + \textcolor{BrickRed}{\Lb\La}\La\Lb \\
        &
        \uncover<4->{=
        2 \La\Lb\Lb\La + \La\Lb\La\Lb + \Lb\La\Lb\La + 2 \Lb\La\La\Lb}
    \end{split}\end{equation*}
    \vspace{-1.5em}
\end{block}}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algebraic issues}

\begin{block}{Proposition}
    The set of binary words of length $n$ that are square w.r.t.
    $\shuffle$ is in one-to-one correspondence with the set of
    square permutations of length $n$ avoiding the patterns
    $213$ and~$231$.
\end{block}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Algorithmic issues}

@@
